# Reproducible Research: Peer Assessment 1

## Preparation
First let's set the locale to English then load the lattice plotting library.
```{r prepare, echo = TRUE}
Sys.setlocale("LC_TIME", "English")

library(lattice)
```

## Loading and preprocessing the data
Now we unzip the data and read it.
```{r load, echo = TRUE}
unzip("activity.zip")

data<-read.csv("activity.csv",stringsAsFactors=FALSE)
```


## What is mean total number of steps taken per day?
That is given by
```{r steps, echo = TRUE}
sd <- aggregate(steps~date, data, "sum")

hist(sd$steps, col="grey", xlab = "steps", main = "Steps")
```
With Mean
```{r mean, echo = TRUE}
mean(sd$steps)
```
and Median
```{r median, echo = TRUE}
median(sd$steps)
```

## What is the average daily activity pattern?
```{r pattern, echo = TRUE}
si <- aggregate(steps~interval, data, "mean")

plot(si, type = "l", main = "Average Steps by Interval", xlab = "Interval", ylab = "Step Avg.")
```
The interval with the greatest average number of steps is
```{r maximum, echo = TRUE}
si[which(si$steps == max(si$steps)),"interval"]
```

With
```{r m_quantity, echo = TRUE}
si[which(si$steps == max(si$steps)),"steps"]
```
steps on average
## Imputing missing values
Total number of missing values is
```{r missing, echo = TRUE}
sum(!complete.cases(data))
```

We'll replace the missing values with the mean of each observation's interval, creating a copy of our data first.
```{r impute, echo = TRUE}
newData<-data
newData[is.na(newData),"steps"] <- si[ match(newData[is.na(newData),"interval"], si$interval) ,"steps"]
hist(newData$steps, col = "grey", main = "Imputed Data. Histogram of steps per day", xlab = "Steps")
```
With Mean
```{r imp_mean, echo = TRUE}
mean(newData$steps)
```
And Median
```{r imp_median, echo = TRUE}
median(newData$steps)
```

No difference , as expected, since we're replacing each NA with the mean.
## Are there differences in activity patterns between weekdays and weekends?
```{r week, echo = TRUE}
newData$day <- as.factor(weekdays( strptime( newData$date, "%Y-%m-%d")))

newData$week <- as.factor(sapply(newData$day, function(d) if (d =="Saturday" || d =="Sunday") "weekend" else "weekday"))

summary <- aggregate(steps ~ interval + week, newData, "mean")

xyplot(steps~interval|week, summary, layout=c(1,2),type="l")
```
Clearly, yes.